---
title: "STA130 Final Project"
author: "Haiyue Yang, Yijing Chen, Xi Zheng(TUT0103-04)"
subtitle: How the launch of request expiry changed user behavior
output:
  beamer_presentation: default
  slidy_presentation: default
  ioslides_presentation: default
  widescreen: yes
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
library(lubridate)
library(readxl)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# This code chunk will run but not appear in the knit document
# Here is a sample showing you how to read in the data
requests <- read_excel("requests_Mar18_2019.xlsx")
categories <- read_excel("data/categories.xlsx")
courses <- read_excel("data/courses.xlsx")
educatoraccounts <- read_excel("data/educatoraccounts.xlsx")
employeraccounts <- read_excel("data/employeraccounts.xlsx")
projects <- read_excel("data/projects.xlsx")
studentaccounts <- read_excel("data/studentaccounts.xlsx")
teams <- read_excel("data/teams.xlsx")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
request_type <- requests %>% 
  mutate(type=ifelse(`Actor Id` %in% educatoraccounts$Id,educatoraccounts$Type,ifelse(`Actor Id` %in% employeraccounts$Id,employeraccounts$Type,'student'))) %>%
  mutate(recipient_type=ifelse(`Recipient Id`%in% educatoraccounts$Id,educatoraccounts$Type,ifelse(`Actor Id` %in% employeraccounts$Id,employeraccounts$Type,'student'))) %>%
  mutate(`Day of Created At`=as.Date(`Day of Created At`, "%d %B, %Y")) %>%
  mutate(`Day of Updated At`=as.Date(`Day of Updated At`, "%d %B, %Y")) %>%
  mutate(`Day of Expired At`=as.Date(`Day of Expired At`, "%d %B, %Y")) %>%
  mutate(`Month of Created At`=substr(`Day of Created At`,1,7)) %>%
  filter(`Day of Created At`!='2018-04-12'&`Day of Created At`!='2018-08-30')

# filter out 2018-4-12 & 2018-8-30 because of extreme data
```


## Introduction

On November 21, 2018, Riipen, a Vancouver-based education technology company, launched a new feature called "request expiry".\
This project is going to show how user behaviour have changed after the launch of request expiry from the following perspectives. 


## Objectives

### Change in the behaviour of recipients
- Whether the **response rate** (in 14days) changed?
- How the **response time** has changed?

### Change in the behaviour of creators
- How the **number of requests created per day** has changed?
- How the **number of accounts created per month** has changed?

## Data Summary

- mutate()&ifelse(): add the type of users into request.xlsx
- mutate()&as.Date(): change variable types from chr to date
- mutate()&substr(): add the month of each request created at
- filter(): filter out data on 2018-4-12 & 2018-8-30 (extreme)

## Response rate (in 14 Days)
### State hypothesis
**Null hypothesis**: $H_0: rate\_diff=0$\
The response rate did not change after the November 21 launch of request expiry.\
**Alternative hypothesis**: $H_a: rate\_diff\ne0$ \
The response rate has changed after the November 21 launch of request expiry.

### Calculate test statistic: 
- Filter out the requests still pending
- Summarise the response rate before and after 2018-11-21
- Calculate the difference.
- Test statistic (rate_diff):\
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
# response rate in 14 days
response_rate <- request_type %>%
  filter(State!='pending') %>%
  mutate(response_time=`Day of Updated At`-`Day of Created At`) %>%
  mutate(rate=ifelse(response_time <= 14,1,0))
rate_before <- response_rate %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
rate_after <- response_rate %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
rate_diff <- rate_after - rate_before
rate_diff
```


## Response rate (in 14 Days)
### Simulate under the null hypothesis
- Shuffle the variable of the day requests created at
- Calculate the difference between the response rate before and after 2018-11-21
- Repeat 1000 times
- Turn the result into a data frame and plot the data

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
set.seed(130)
rep <- 1000
sim_rate <- rep(NA,rep)
for (i in 1:rep)
{
  response_rate_ <- response_rate %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  rate_before_ <- response_rate_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  rate_after_ <- response_rate_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  sim_rate[i] <- rate_after_ - rate_before_
}
sim_rate <- tibble(sim_rate)
sim_rate %>% ggplot(aes(x=sim_rate)) + 
  geom_histogram() + 
  geom_vline(xintercept = rate_diff, color="red") + 
  geom_vline(xintercept=-rate_diff, color="red") +
  labs(x='difference in response rate') +
  ggtitle(label = "simulated sampling distribution of difference in response rate")
```


## Response rate (in 14 Days)
### Assess evidence against the null hypothesis
- Compare the test statistic to the simulated sampling distribution
- Calculate the p-value:
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
extreme_rate <- sim_rate %>%
  filter(sim_rate>=rate_diff | sim_rate<=-rate_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_value_rate <- extreme_rate/rep
p_value_rate # p-value < 0.05
```
### Conclusion
p-value=0.009<0.01\
We have strong evidence against the null hypothesis

## Response rate (in 14 Days)
### **Different types of recipients**
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
response_stu <- response_rate %>% filter(recipient_type=="student")
response_emp <- response_rate %>% filter(recipient_type=="employer")
```
**professor**: too few observations to draw a conclusion\
**student**:

- test statistic: stu_diff=0.1768805
```{r, echo=FALSE, message=FALSE, warning=FALSE}
stu_rate_before <- response_stu %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
stu_rate_after <- response_stu %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
stu_diff <- stu_rate_after - stu_rate_before
```
- p-value=0
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# hypothesis test
# H0 stu_diff=0
set.seed(130)
sim_stu <- rep(NA,rep)
for (i in 1:rep)
{
  stu_ <- response_stu %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  s_before_ <- stu_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  s_after_ <- stu_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  sim_stu[i] <- s_after_ - s_before_
}
sim_stu <- tibble(sim_stu)
extreme_s <- sim_stu %>%
  filter(sim_stu >=stu_diff | sim_stu <=-stu_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_stu <- extreme_s/rep
```
- 95% confidence interval
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# bootstrap
set.seed(130)
boot_stu <- rep(NA,rep)
ns <- response_stu %>% summarise(n()) %>% as.numeric()
for (i in 1:rep)
{
  bs <- response_stu %>% sample_n(size=ns,replace=TRUE)
  bs_before <- bs %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  bs_after <- bs %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  boot_stu[i] <- bs_after-bs_before
}
boot_stu %>% quantile(c(0.025,0.975))
```

**employer**: 

- test statistic: emp_diff=0.01536152
```{r, echo=FALSE, message=FALSE, warning=FALSE}
emp_rate_before <- response_emp %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
emp_rate_after <- response_emp %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
emp_diff <- emp_rate_after - emp_rate_before
```
- p-value=0.706>0.1
```{r, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(130)
sim_emp <- rep(NA,rep)
for (i in 1:rep)
{
  emp_ <- response_emp %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  s_before_ <- emp_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  s_after_ <- emp_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  sim_emp[i] <- s_after_ - s_before_
}
sim_emp <- tibble(sim_emp)
extreme_s <- sim_emp %>%
  filter(sim_emp >=emp_diff | sim_emp <=-emp_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_emp <- extreme_s/rep# p-value>0.1
```

## Response time
### **hypothesis test & p-value**
- hypothesis- hypothesis\
**Null hypothesis**: \
The average response time did not change after the November 21 launch of request expiry.
$$H_0: time\_diff=0$$
**Alternative hypothesis**: \
The average response time has changed after the November 21 launch of request expiry.
$$H_a: time\_diff\ne0$$
- test statistic: (time_diff)\
```{r,echo=FALSE, message=FALSE, warning=FALSE }
response <- request_type %>%
  filter(`State`!='expired') %>%
  filter(`State`!='pending') %>%
  filter(`Day of Created At` != '2018-04-12') %>%
  filter(`Day of Created At` != '2018-08-30')
response$response_time <- response$`Day of Updated At` - response$`Day of Created At`

# average response time each day
time <- response %>%
  group_by(`Day of Created At`) %>%
  summarise(time_ave_each_day=mean(response_time))

time_before <- time %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean=mean(time_ave_each_day)) %>%
  as.numeric()

time_after <- time %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean=mean(time_ave_each_day)) %>%
  as.numeric()

time_diff <- time_after - time_before
time_diff
```

## Response time
### **hypothesis test & p-value**
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
set.seed(130)
sim_time <- rep(NA,rep)
for (i in 1:rep)
{
  time_ <- time %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  time_before_ <- time_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(time_ave_each_day)) %>%
    as.numeric()
  time_after_ <- time_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(time_ave_each_day)) %>%
    as.numeric()
  sim_time[i] <- time_after_-time_before_
}
sim_time <- tibble(sim_time)
sim_time %>% ggplot(aes(x=sim_time)) + 
  geom_histogram() + 
  geom_vline(xintercept = time_diff, color="red") + 
  geom_vline(xintercept=-time_diff, color="red") +
  labs(x='difference in average response time') +
  ggtitle(label="Simulated Sampling Distribution of the difference in average responce time")
```

- p-value
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
extreme_time <- sim_time %>%
  filter(sim_time>= abs(time_diff) | sim_time<= -abs(time_diff)) %>%
  summarise(n()) %>%
  as.numeric()
p_value_time <- extreme_time/rep
p_value_time
```
- conclusion\
p-value=0\
We have strong evidence against the null hypothesis.

## Response time
### **bootstrap sampling & confidence interval**
### bootstrap sample
- Draw 1000 bootstrap samples of the same size as the original sample with replacement
- For each bootstrap sample, calculate the statistic
- Summarise the value of statistic for all bootstrap samples 
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
set.seed(130)
boot_time <- rep(NA,rep)
st <- time %>% summarise(n()) %>% as.numeric()
for (i in 1:rep)
{
  time2 <- time %>% sample_n(size=st,replace=TRUE)
  time_before2 <- time2 %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(time_ave_each_day)) %>%
    as.numeric()
  time_after2 <- time2 %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(time_ave_each_day)) %>%
    as.numeric()
  boot_time[i] <- time_after2-time_before2
}
bt1 <- boot_time %>% quantile(0.025) %>% as.numeric()
bt2 <- boot_time %>% quantile(0.975) %>% as.numeric()
boot_time <- tibble(boot_time)
boot_time %>% ggplot(aes(x=boot_time)) + 
  geom_histogram() +
  geom_vline(aes(xintercept=bt1),color="blue") +
  geom_vline(aes(xintercept=bt2),color="blue") +
  labs(x="sample difference in average response time") +
  ggtitle(label="bootstrap sampling distribution of the difference in avverage responce time")
```

## Response time
### **bootstrap sampling & confidence interval**
### 95% confidence interval
- Calculate the middle 95% interval of values of the bootstrap statistic
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
boot_time$boot_time %>% quantile(c(0.025,0.975)) # 95% confidence interva
```
We are 95% confident that the response time has decreased by between 9.470533 and -17.154014 days after 2018-11-21.


## Number of requests per day
### **clear out outliers**
We plotted a scartter plot for number of requests created each day.\
From the plot, we discovered an outlier at 2019-02-28 and cleared it out.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5}
# request each day
num <- request_type %>%
  group_by(`Day of Created At`) %>%
  summarise(num=n())
# discover an outlier at 2019-02-28
num %>% ggplot(aes(x=`Day of Created At`,y=num)) + 
  geom_point() + 
  geom_point(aes(x=as.Date('2019-02-28'),y=51),color="red") +
  labs(y='number of requests created') +
  ggtitle(label="Number of Requests Created Each Day")
num <- num %>% filter(`Day of Created At`!='2019-02-28') # filter out the outlier
```

## Number of requests per day
### **hypothesis test & p-value**
- hypothesis\
**Null hypothesis**: \
The number of requests created per day did not change after the November 21 launch of request expiry.
$$H_0: num\_diff=0$$
**Alternative hypothesis**: \
The number of requests created per day has changed after the November 21 launch of request expiry.
$$H_a: num\_diff\ne0$$
- test statistic: (num_diff)\
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
num_before <- num %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean=mean(num)) %>%
  as.numeric()
num_after <- num %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean=mean(num)) %>%
  as.numeric()
num_diff <- num_after - num_before
num_diff
```

## Number of requests per day
### **hypothesis test & p-value**
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=4}
# hypothesis type
# H0: difference=0
set.seed(130)
sim <- rep(NA,rep)
for (i in 1:rep)
{
  num_ <- num %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  num_before_ <- num_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  num_after_ <- num_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  sim[i] <- num_after_-num_before_
}
sim <- tibble(sim)
sim %>% ggplot(aes(x=sim)) + 
  geom_histogram() + 
  geom_vline(xintercept = num_diff, color="red") + 
  geom_vline(xintercept=-num_diff, color="red") +
  labs(x='difference in average number of request each day') +
  ggtitle(label="Simulated difference in average number of request created each day")
```

- p-value
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
extreme <- sim %>%
  filter(sim>=num_diff | sim<=-num_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_value <- extreme/rep
p_value # p-value < 0.05
```

- conclusion\
p-value=0.024<0.05\
We have evidence against the null hypothesis.


## Number of requests per day
### **bootstrap sampling & confidence interval**
- bootstrap sampling distribution:
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
# bootstrap together
set.seed(130)
boot1 <- rep(NA,rep)
s <- num %>% summarise(n()) %>% as.numeric()
for (i in 1:rep)
{
  num2 <- num %>% sample_n(size=s,replace=TRUE)
  num_before2 <- num2 %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  num_after2 <- num2 %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  boot1[i] <- num_after2-num_before2
}
b1 <- boot1 %>% quantile(0.025) %>% as.numeric()
b2 <- boot1 %>% quantile(0.975) %>% as.numeric()
boot1 <- tibble(boot1)
boot1 %>% ggplot(aes(x=boot1)) + 
  geom_histogram() +
  geom_vline(aes(xintercept=b1),color="blue") +
  geom_vline(aes(xintercept=b2),color="blue") +
  labs(x="difference in average number of request per day") +
  ggtitle(label="bootstrap sampling distribution of difference in average number of request per day")
```
- 95% confidence interval
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
boot1$boot1 %>% quantile(c(0.025,0.975)) # 95% confidence interval
```

## Number of new accounts created
### **clean up data**
- summarise the number of accounts craeted each month
- filter out accounts created on '2017-02-11' because it is the start date.
- filter out accounts created in '2019-03' because the month is not complete.
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
studentaccounts <- studentaccounts %>%
  mutate(`Day of Created At`=as.Date(`Day of Created At`, "%d %B, %Y")) %>%
  mutate(`Month of Created At`=substr(`Day of Created At`,1,7))
employeraccounts <- employeraccounts %>%
  mutate(`Day of Created At`=as.Date(`Day of Created At`, "%d %B, %Y")) %>%
  mutate(`Month of Created At`=substr(`Day of Created At`,1,7))
educatoraccounts <- educatoraccounts %>%
  mutate(`Day of Created At`=as.Date(`Day of Created At`, "%d %B, %Y")) %>%
  mutate(`Month of Created At`=substr(`Day of Created At`,1,7))
stu_creat <- studentaccounts %>%
  filter(`Day of Created At`!='2017-02-11') %>%
  group_by(`Month of Created At`) %>%
  summarise(stu=n())
emp_creat <- employeraccounts %>%
  filter(`Day of Created At`!='2017-02-11') %>%
  group_by(`Month of Created At`) %>%
  summarise(emp=n())
edu_creat <- educatoraccounts %>%
  filter(`Day of Created At`!='2017-02-11') %>%
  group_by(`Month of Created At`) %>%
  summarise(edu=n())
merge1 <- merge(stu_creat,emp_creat,all=TRUE,sort=TRUE)
account <- merge(merge1, edu_creat, all=TRUE,sort=TRUE) %>%
  mutate(all=stu+emp+edu) %>%
  filter(`Month of Created At`!='2019-03')
```

## Number of new accounts created
### **hypothesis test & p-value**

- hypothesis\
**Null hypothesis**: \
The number of accounts created each month did not change after the November 21 launch of request expiry.
$$H_0: account\_diff=0$$
**Alternative hypothesis**: \
The number of accounts created each month has changed after the November 21 launch of request expiry.
$$H_a: account\_diff\ne0$$
- test statistic (account_diff):\
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
account_before <- account %>%
  filter(`Month of Created At`!='2018-12'&`Month of Created At`!='2019-01'&`Month of Created At`!='2019-02'&`Month of Created At`!='2019-03') %>%
  summarise(mean(all)) %>%
  as.numeric()
account_after <-account %>%
  filter(`Month of Created At`=='2018-12'|`Month of Created At`=='2019-01'|`Month of Created At`=='2019-02') %>% 
  summarise(mean(all)) %>%
  as.numeric() #does not inclue'2019-03'
account_diff <- account_after-account_before
account_diff
```

## Number of new accounts created
### **hypothesis test & p-value**
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
set.seed(130)
sim_act <- rep(NA,rep)
for (i in 1:rep)
{
  account_ <- account %>% mutate(`Month of Created At`=sample(`Month of Created At`))
  account_before_ <- account_ %>%
    filter(`Month of Created At`!='2018-12'&`Month of Created At`!='2019-01'&`Month of Created At`!='2019-02'&`Month of Created At`!='2019-03') %>%
    summarise(mean(all)) %>%
    as.numeric()
  account_after_ <-account_ %>%
    filter(`Month of Created At`=='2018-12'|`Month of Created At`=='2019-01'|`Month of Created At`=='2019-02') %>% 
    summarise(mean(all)) %>%
    as.numeric() #does not inclue'2019-03'
  sim_act[i] <- account_after_-account_before_
}
sim_act <- tibble(sim_act)
sim_act %>% ggplot(aes(x=sim_act)) + 
  geom_histogram() + 
  geom_vline(xintercept = account_diff, color="red") + 
  geom_vline(xintercept=-account_diff, color="red") +
  labs(x='simulated difference in average number of new account created each month')
```
- p-value
```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
extreme_act <- sim_act %>%
  filter(sim_act >=account_diff | sim_act <=-account_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_act <- extreme_act/rep
p_act
```
- conclusion
p-value=0.476>0.1
We do not have evidence against the null hypothesis.

## Conclusion
### Change in Behaviour of Recipients
**response rate**

- The response rate has changed after the launch of request expiry.
- More specifically:\
the response rate of employers did not change;
the response rate of students has changed, and we are 95% confident that it is increased by between 0.07873107 and 0.27436534.

**response time**

- The average response time has changed after the launch of request expiry.
- We are 95% confident that the average responce time has decreased by between 9.470533 and 17.154014 days.


## Conclusion
### Change in Behaviour of Creators
**number of requests each day**

- The number of requests created each day has changed after the launch of request expiry.
- We are 95% confident that the average number of requests per day has increased by between 0.1232033 and 1.7172502

**number of new accounts each month**

- The number of new accounts created each month did not change after the launch of request expiry.

## Conclusion
### possible explanations

- The launch of request expiry has increased the response rate within 14 days and decreased the responses time, and thus improved user experience.
- Thus,users tend to use Riipen more.
- Since request expiry has launched only 4 months, it has not draw more new users, but is probably going to draw more users.

**limitation**

We did not analyze how many requests are recent after expiry, so probably the number of requests created increase simply because some requests are recent after expiry.




