---
title: "STA130 Final Poster Project"
author: "Haiyue Yang  Xi Zheng  Yijing Chen, TUT0103, 04"
subtitle: The subtitle of my project
output:
  beamer_presentation: default
  ioslides_presentation: default
  widescreen: yes
---


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# echo=FALSE will stop the code chunk from appearing in the knit document
# warning=FALSE and message=FALSE will stop R messages from appearing in the knit document
library(tidyverse)
library(lubridate)
library(readxl)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# This code chunk will run but not appear in the knit document
# Here is a sample showing you how to read in the data
requests <- read_excel("requests_Mar18_2019.xlsx")
categories <- read_excel("data/categories.xlsx")
courses <- read_excel("data/courses.xlsx")
educatoraccounts <- read_excel("data/educatoraccounts.xlsx")
employeraccounts <- read_excel("data/employeraccounts.xlsx")
projects <- read_excel("data/projects.xlsx")
studentaccounts <- read_excel("data/studentaccounts.xlsx")
teams <- read_excel("data/teams.xlsx")
```


```{r}
request_type <- requests %>% 
  mutate(type=ifelse(`Actor Id` %in% educatoraccounts$Id,educatoraccounts$Type,ifelse(`Actor Id` %in% employeraccounts$Id,employeraccounts$Type,'student'))) %>%
  mutate(recipient_type=ifelse(`Recipient Id`%in% educatoraccounts$Id,educatoraccounts$Type,ifelse(`Actor Id` %in% employeraccounts$Id,employeraccounts$Type,'student'))) %>%
  mutate(`Day of Created At`=as.Date(`Day of Created At`, "%d %B, %Y")) %>%
  mutate(`Day of Updated At`=as.Date(`Day of Updated At`, "%d %B, %Y")) %>%
  mutate(`Day of Expired At`=as.Date(`Day of Expired At`, "%d %B, %Y")) %>%
  mutate(`Month of Created At`=substr(`Day of Created At`,1,7)) %>%
  filter(`Day of Created At`!='2018-04-12'&`Day of Created At`!='2018-08-30')

# filter out 2018-4-12 & 2018-8-30 because of extreme data

request_type
```

```{r}
# request each day
num <- request_type %>%
  group_by(`Day of Created At`) %>%
  summarise(num=n())
# discover an outlier at 2019-02-28
num %>% ggplot(aes(x=`Day of Created At`,y=num)) + 
  geom_point() + 
  geom_point(aes(x=as.Date('2019-02-28'),y=51),color="red") +
  labs(y='number of requests each day')
num <- num %>% filter(`Day of Created At`!='2019-02-28') # filter out the outlier

num_before <- num %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean=mean(num)) %>%
  as.numeric()
num_after <- num %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean=mean(num)) %>%
  as.numeric()
num_diff <- num_after - num_before
num_diff

# hypothesis type
# H0: difference=0
set.seed(130)
rep <- 1000
sim <- rep(NA,rep)
for (i in 1:rep)
{
  num_ <- num %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  num_before_ <- num_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  num_after_ <- num_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  sim[i] <- num_after_-num_before_
}
sim <- tibble(sim)
sim %>% ggplot(aes(x=sim)) + 
  geom_histogram() + 
  geom_vline(xintercept = num_diff, color="red") + 
  geom_vline(xintercept=-num_diff, color="red") +
  labs(x='simulated difference in average number of request each day')
extreme <- sim %>%
  filter(sim>=num_diff | sim<=-num_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_value <- extreme/rep
p_value # p-value < 0.05
```

```{r}
# bootstrap together
set.seed(130)
boot1 <- rep(NA,rep)
s <- num %>% summarise(n()) %>% as.numeric()
for (i in 1:rep)
{
  num2 <- num %>% sample_n(size=s,replace=TRUE)
  num_before2 <- num2 %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  num_after2 <- num2 %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean=mean(num)) %>%
    as.numeric()
  boot1[i] <- num_after2-num_before2
}
b1 <- boot1 %>% quantile(0.025) %>% as.numeric()
b2 <- boot1 %>% quantile(0.975) %>% as.numeric()
boot1 <- tibble(boot1)
boot1 %>% ggplot(aes(x=boot1)) + 
  geom_histogram() +
  geom_vline(aes(xintercept=b1),color="blue") +
  geom_vline(aes(xintercept=b2),color="blue") +
  labs(x="sample difference in average request from bootstrap samples")
boot1$boot1 %>% quantile(c(0.025,0.975)) # 95% confidence interval
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
# response rate in 14 days
response_rate <- request_type %>%
  filter(State!='pending') %>%
  mutate(response_time=`Day of Updated At`-`Day of Created At`) %>%
  mutate(rate=ifelse(response_time <= as_datetime(14),1,0))
response_rate
rate_before <- response_rate %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
rate_after <- response_rate %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
rate_diff <- rate_after - rate_before
rate_diff

# hypothesis test
# H0:rate_diff=0
set.seed(130)
sim_rate <- rep(NA,rep)
for (i in 1:rep)
{
  response_rate_ <- response_rate %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  rate_before_ <- response_rate_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  rate_after_ <- response_rate_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  sim_rate[i] <- rate_after_ - rate_before_
}
sim_rate <- tibble(sim_rate)
sim_rate %>% ggplot(aes(x=sim_rate)) + 
  geom_histogram() + 
  geom_vline(xintercept = rate_diff, color="red") + 
  geom_vline(xintercept=-rate_diff, color="red") +
  labs(x='simulated difference in response rate')
extreme_rate <- sim_rate %>%
  filter(sim_rate>=rate_diff | sim_rate<=-rate_diff) %>%
  summarise(n()) %>%
  as.numeric()
p_value_rate <- extreme_rate/rep
p_value_rate # p-value < 0.05
```

```{r}
response_rate %>% filter(recipient_type=="student")
response_rate %>% filter(recipient_type=="employer")
response_rate %>% filter(recipient_type=="professor") # observations of professors are too few to draw a conclusion
response_stu <- response_rate %>% filter(recipient_type=="student")
response_emp <- response_rate %>% filter(recipient_type=="employer")
stu_rate_before <- response_stu %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
stu_rate_after <- response_stu %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
stu_diff <- stu_rate_after - stu_rate_before
emp_rate_before <- response_emp %>%
  filter(`Day of Created At`<'2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
emp_rate_after <- response_emp %>%
  filter(`Day of Created At`>='2018-11-21') %>%
  summarise(mean(rate)) %>%
  as.numeric()
emp_diff <- emp_rate_after - emp_rate_before
type_diff <- emp_diff - stu_diff
type_diff

# hypothesis test 
# H0: type_diff=0
set.seed(130)
sim_type <- rep(NA,rep)
for (i in 1:rep)
{
  type <- response_rate %>% 
    filter(recipient_type!="professor") %>% 
    mutate(recipient_type=sample(recipient_type))
  stu_before <- type %>%
    filter(recipient_type=="student") %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  stu_after <- type %>%
    filter(recipient_type=="student") %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  stu <- stu_after-stu_before
  emp_before <- type %>%
    filter(recipient_type=="employer") %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  emp_after <- type %>%
    filter(recipient_type=="employer") %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  emp <- emp_after-emp_before
  sim_type[i] <-emp-stu
}
sim_type <- tibble(sim_type)
sim_type %>% ggplot(aes(x=sim_type)) + 
  geom_histogram() + 
  geom_vline(xintercept = type_diff, color="red") + 
  geom_vline(xintercept=-type_diff, color="red") +
  labs(x='simulated difference in difference in response rate')
extreme_type <- sim_type %>%
  filter(sim_type>=abs(type_diff) | sim_type<=-abs(type_diff)) %>%
  summarise(n()) %>%
  as.numeric()
p_value_type <- extreme_type/rep
p_value_type # p-value < 0.05
```

```{r}
# student_difference
response_stu
stu_diff

# hypothesis test
# H0 stu_diff=0
sim_stu <- rep(NA,rep)
for (i in 1:rep)
{
  stu_ <- response_stu %>% mutate(`Day of Created At`=sample(`Day of Created At`))
  s_before_ <- stu_ %>%
    filter(`Day of Created At`<'2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  s_after_ <- stu_ %>%
    filter(`Day of Created At`>='2018-11-21') %>%
    summarise(mean(rate)) %>%
    as.numeric()
  sim_stu[i] <- s_after_ - s_before_
}
```

# A Section Heading on its own page

## A heading that starts on a new page

This document shows a few basics on making slides with R markdown.

To produce the slides, Knit to HMTL (ioslides).

(You can also knit to PDF if you have LaTeX installed on your computer.)


---

Three dashes starts a new page when it is not started by a new header.

---

### This is a heading that doesn't start a new page

You can make text **bold** or in *italics*.

To make bullet points, start the points after a blank line:
 
- point one
- another point


---

Include some R output

```{r}
ggplot(cars, aes(x=speed, y=dist)) + geom_point()
```

---

Include the output without showing the code and R messages (which is what you want for your poster presentation).  This R code chunk also changes the size of the plot.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3}
ggplot(cars, aes(x=speed, y=dist)) + geom_point()
```

This plot shows that there is a positive relationship between distance and speed.

## Read more

There is more information on R markdown on the course website at [https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html](https://ntaback.github.io/UofT_STA130/Rmarkdownforclassreports.html)

For more on creating an `ioslides` presentation see [https://rmarkdown.rstudio.com/ioslides_presentation_format.html](https://rmarkdown.rstudio.com/ioslides_presentation_format.html)

# Headings you should include in your project

## Introduction

Include here a few sentences to introduce the problem and provide context. You might want to briefly summarize the data in words (what is the data and what is it used for). You can present the questions you are investigating here. 


## Objectives (optional)

You can list the questions of interest in complete English sentences here to highlight them. 

## Data Summary (optional)

Here you can explain how you cleaned the data and created variables suitable for answering your questions. 
You can also include graphical displays that either motivated or address the questions.

## Statistical Methods

Describe here what you have done to the data without presenting any results (output). If you want to indicate variables by symbols or variable names, define them here. 

## Results

Present the main results here, in order of importance, related to the questions asked. You might use tables or graphs, or other ways to summarize your results.

## Conclusion

Give your main conclusions here. Follow the order of questions you presented. 

Here you can also mention any additional considerations, concerns, or issues you might have. For example, if the results were unexpected, you can discuss this and perhaps offer possible explanations.

## Acknowledgements (optional)

If you received any help from someone other than your team members you can acknowledge them. For example:   
*The authors thank Wei for providing information on additional data resources. The authors would like to thank "TA name" for helpful suggestions and comments that improved the presentation of this poster.*



